Range Analysis of Whole Programs

### 2012-03-29 doc/Manuscript/paper.tex
Changed the paper, and added the pdf to the svn.

\noindent
\textbf{The impact of strong components in time and precision.}
Figure~\ref{fig:impactEssaSCC}(a) shows that strong components dramatically
decrease the runtime of our analysis.
For instance, this design improves the analysis time of \texttt{483.xalancbmk}
452 times!
Moreover, as the example in Section~\ref{sub:micro} shows, our algorithm benefits
from the partitioning of constraint graphs into SCCs to be more precise.
Figure~\ref{fig:impactEssaSCC}(b) measures this precision.
Range propagation along strong components allows us, for example, to reduce
41.5\% more bits in \texttt{401.bzip}.

@@

\noindent
\textbf{The impact of whole program analysis on precision.}
Figure~\ref{fig:essaDivSSA}(Top) shows that the whole program analysis increases
moderately the precision of our analysis.
We show results for the five largest programs in three different categories of
benchmarks: SPEC CPU 2006, the Stanford Suite~\footnote{\texttt{http://classes.engineering.wustl.edu/cse465/docs/BCCExamples/stanford.c}} and
Bitwise~\cite{Stephenson00}.
Our initial goal when developing this analysis was to support a bitwidth-aware
register allocator.
Thus, we measure precision by the average number of bits that our
analysis allows us to save per program variables.
It is very important to notice that we do not consider the constants in the
program text in our statics of precision.
In other words, we only measure bitwidth reduction in variables that a constant
propagation step could not remove.
Our results for the SPEC programs were disappointingly: on the average, the
intra-procedural version of our analysis saves 5.23\% of the total program
bitwidth.
By using the inter-procedural version, with function inlining to simulate
context-sensitiveness, we can increase this number to 8.89\%.
A manual inspection of the SPEC programs reveal that this result is expected:
they manipulate files, and their code do no provide enough
explicit constants to power our analysis up.
However, with numerical benchmarks we fare much better.
On the average our inter-procedural algorithm reduces the bitwidth of the
Stanford benchmarks by 36.24\%.
Notice that this number is better than the one we get with inlining (31.97\%),
because functions in those benchmarks are called only at one location.
Finally, for Bitwise we obtain a bitwidth reduction of 12.27\%.
However, this average is brought down by two outliers: \texttt{edge\_detect} and
\texttt{sha}, which cannot be reduced.
The bitwise benchmarks were implemented by Stephenson
{\em et al.}~\cite{Stephenson00} to validate their bitwidth analysis.
Our results are on par with those found by the original authors.
The bitwise programs contain only the \texttt{main} function; thus, different
versions of our algorithm find the same results.

@@

Our implementation of the range analysis algorithm described in this paper is
publicly available at \texttt{http://google.code/range-analysis/}.
This repository contains instructions about how to deploy and use our
implementation.
We provide a gallery of examples, including source codes,
CFGs and constraint graphs that we produce for meaningful programs at
\texttt{http://google.code/range-analysis/wiki/gallery}.

>>>>

\noindent
\textbf{The impact of strong components in time and precision.}
Figure~\ref{fig:impactEssaSCC}(a) shows that strong components dramatically
decrease the runtime of our analysis.
For instance, this design improves the analysis time of \texttt{483.xalancbmk}
452 times!
Moreover, as the example in Section~\ref{sub:micro} shows, our algorithm benefits
from the partitioning of constraint graphs into SCCs to be more precise.
Figure~\ref{fig:impactEssaSCC}(b) measures this precision.
Range propagation along strong components allows us, for example, to reduce
41.5\% more bits in \texttt{401.bzip2}.

@@

\noindent
\textbf{The impact of whole program analysis on precision.}
Figure~\ref{fig:essaDivSSA}(Top) shows that the whole program analysis increases
moderately the precision of our analysis.
We show results for the five largest programs in three different categories of
benchmarks: SPEC CPU 2006, the Stanford Suite~\footnote{\texttt{http://classes.engineering.wustl.edu/cse465/docs/BCCExamples/stanford.c}} and
Bitwise~\cite{Stephenson00}.
Our initial goal when developing this analysis was to support a bitwidth-aware
register allocator.
Thus, we measure precision by the average number of bits that our
analysis allows us to save per program variables.
It is very important to notice that we do not consider constants in our statistics of precision.
In other words, we only measure bitwidth reduction in variables that a constant
propagation step could not remove.
Our results for the SPEC programs were disappointingly: on the average, the
intra-procedural version of our analysis saves 5.23\% of the total program
bitwidth.
By using the inter-procedural version, with function inlining to simulate
context-sensitiveness, we can increase this number to 8.89\%.
A manual inspection of the SPEC programs reveal that this result is expected:
they manipulate files, and their code do no provide enough
explicit constants to power our analysis up.
However, with numerical benchmarks we fare much better.
On the average our inter-procedural algorithm reduces the bitwidth of the
Stanford benchmarks by 36.24\%.
Notice that this number is better than the one we get with inlining (31.97\%),
because functions in those benchmarks are called only at one location.
Finally, for Bitwise we obtain a bitwidth reduction of 12.27\%.
However, this average is brought down by two outliers: \texttt{edge\_detect} and
\texttt{sha}, which cannot be reduced.
The bitwise benchmarks were implemented by Stephenson
{\em et al.}~\cite{Stephenson00} to validate their bitwidth analysis.
Our results are on par with those found by the original authors.
The bitwise programs contain only the \texttt{main} function; thus, different
versions of our algorithm find the same results.

@@

Our implementation of the range analysis algorithm described in this paper is
publicly available at \texttt{http://code.google.com/p/range-analysis/}.
This repository contains instructions about how to deploy and use our
implementation.
We provide a gallery of examples, including source codes,
CFGs and constraint graphs that we produce for meaningful programs at
\texttt{http://code.google.com/p/range-analysis/wiki/gallery}.
